---
title: "CruiseCheck"
author: "Dave Warner"
date: "11/13/2019"
output: html_document
---
### This code will provide you an overview of the characteristics of trawl data tables. It will only work for csv data
if your files are called op, tr_op, tr_catch, tr_l, tr_lf, tr_fish. It will fail with any other file names. It is just that simple. Fix the file names or edit the code to accept a wider range of file names with can be coerced to these and I will
gladly add that to the package/function.

# <u>Op Data Characteristics</u>
The first step in examing csv files is to see if they have the correct field names. If they do not, the user will be prompted to remedy this with a message stating which field names are not correct. This step is only necessary for csv files.


```{r echo=FALSE, message = FALSE, warning = FALSE}
library(janitor)
library(tidyverse)
library(CruiseCheckeR)

data("op_names")
op <- op %>% clean_names(case = "all_caps")
  obs_names <- names(op)
  #check for required fields that can't be NULL. Excepted for csv inputs
  #is 'OP_ID'
required <- c('YEAR', 'VESSEL', 'SERIAL', 'SAMPLE_TYPE', 'LAKE', 'PORT',                             'CRUISE', 'OP_DATE', 'TIME')  
missed <- setdiff(required, obs_names)
if (length(missed) > 0)
    stop(print(paste0("The Op data are missing the required field(s) ", missed)))
```


# <u>Table of Op Variables</u>
```{r echo=FALSE, message = FALSE, warning = FALSE}
df <- as.data.frame(CruiseCheckeR::dfSmry(op))

knitr::kable(df)
```

# <u>Plots of Op Variables</u>
```{r echo=FALSE, message = FALSE, warning = FALSE, fig.dim = c(12,10)}
dfPlot(op, mcex=1.1, cex = 0.8, mfrow=c(2,2))
```


# <u>Map of Operation Locations</u>
```{r  echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.dim = c(12,10)}

library(leaflet)
library(leaflet.extras)
library(htmltools)
library(htmlwidgets)
library(mapedit)
library(mapview)

#list of possible location fields
begfields <- c("BEG_LATITUDE_DD",  "BEG_LONGITUDE_DD")
endfields <- c("END_LATITUDE_DD", 
               "END_LONGITUDE_DD")

#if (names(op) %in% c(begfields))
leaflet(op) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~BEG_LONGITUDE_DD, lat = ~BEG_LATITUDE_DD, opacity=0.1, radius =4,
             popup = paste("Serial:", op$SERIAL, "<br>",
                           "Beg_Depth:", op$BEG_DEPTH, "<br>",
                           "Op_Date:", op$OP_DATE, "<br>",
                           "Vessel:", op$VESSEL, "<br>",
                           "Start Latitude:", op$BEG_LATITUDE_DD, "<br>",
                           "Start Longitude:", op$BEG_LONGITUDE_DD))
```

```{r echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.dim = c(12,10)}
#if (names(op) %in% c(endfields))
#leaflet(op) %>%
#  addTiles() %>%
#  addCircleMarkers(lng = ~END_LONGITUDE_DD, lat = ~END_LATITUDE_DD, opacity=0.1, radius =4,
#             popup = paste("Serial:", op$SERIAL, "<br>",
#                           "Beg_Depth:", op$BEG_DEPTH, "<br>",
#                           "Op_Date:", op$OP_DATE, "<br>",
#                           "Vessel:", op$VESSEL, "<br>",
#                           "End Latitude:", op$END_LATITUDE_DD, "<br>",
#                           "End Longitude:", op$END_LONGITUDE_DD))
```


```{r echo=FALSE, message = FALSE, warning = FALSE}
data("tr_op_names")
tr_op <- tr_op %>% clean_names(case = "all_caps")
obs_names <- names(tr_op)
#check for required fields that can't be NULL. Excepted for csv inputs
#is 'OP_ID'

if (dat.source =="csv") 
{
required <- c('YEAR', 'VESSEL', 'SERIAL', 
      "TOW_TIME", "SPEED", "SPEED_UNIT",
      "TYPE_SET", "MESH_SIZE", "TR_DESIGN")
} else {
  required <- c("OP_ID",  
      "TOW_TIME", "SPEED", "SPEED_UNIT",
      "TYPE_SET", "MESH_SIZE", "TR_DESIGN")
}
missed <- setdiff(required, obs_names)
if (length(missed) > 0)
    stop(print(paste0("The tr_op data are missing the required field(s) ", missed)))
```

# <u>Table of TR_OP Variables</u>
```{r echo=FALSE, message = FALSE, warning = FALSE}
df <- as.data.frame(CruiseCheckeR::dfSmry(tr_op))
knitr::kable(df)
```

# <u>Plots of TR_OP Variables</u>
```{r echo=FALSE, message = FALSE, warning = FALSE, fig.dim = c(12,10)}
dfPlot(tr_op, mcex=1.1, cex = 0.8, mfrow=c(2,2))
```



```{r echo=FALSE, message = FALSE, warning = FALSE}
data("tr_catch_names")
tr_catch <- tr_catch %>% clean_names(case = "all_caps")
tr_catch$avg_wt <- tr_catch$WEIGHT/tr_catch$N
obs_names <- names(tr_catch)
#check for required fields that can't be NULL. Excepted for csv inputs
#is 'OP_ID'

if (dat.source =="csv") 
{
required <- c('YEAR', 'VESSEL', 'SERIAL', 
      "SPECIES", "N")
} else {
  required <- c("OP_ID",  "SPECIES", "N")
}
missed <- setdiff(required, obs_names)
if (length(missed) > 0)
    stop(print(paste0("The tr_catch data are missing the required field(s) ", missed)))
```

# <u>Table of TR_CATCH Variables</u>
```{r echo=FALSE, message = FALSE, warning = FALSE}
df <- as.data.frame(CruiseCheckeR::dfSmry(tr_catch))

knitr::kable(df)
```

# <u>Plots of TR_CATCH Variables</u>
```{r echo=FALSE, message = FALSE, warning = FALSE, fig.dim = c(12,10)}
library(plotly)
# These are canned plots for all variables.
dfPlot(tr_catch, mcex=1.1, cex = 0.8, mfrow=c(2,2))
# Now custom plot to look for errors in catch weight
tr_catch$Species <- factor(tr_catch$SPECIES)
options(scipen=10000)
gg <- ggplot()+
  geom_point(data = tr_catch, aes(x = Species, y = avg_wt))+
  scale_y_log10()+
  theme_bw()+
  ggtitle("Plot of Catch Weight/Catch N by Species")+
  labs(y = "Weight/N (g)")

ggplotly(gg)
```

# <u>Table of TR_L Variables</u>
```{r echo=FALSE, message = FALSE, warning = FALSE}
if (dat.source !="csv") {
df <- as.data.frame(CruiseCheckeR::dfSmry(tr_l))
knitr::kable(df)
} else {
  print("The data source is csv, so the TR_L table does not exist!")
}
```

# <u>Plots of TR_L Variables</u>
```{r echo=FALSE, message = FALSE, warning = FALSE, fig.dim = c(12,10)}
if (dat.source !="csv") {
# These are canned plots for all variables.
dfPlot(tr_l, mcex=1.1, cex = 0.8, mfrow=c(2,2))
} else {
    print("The data source is csv, so the TR_L table does not exist!")
}
```  

```{r echo=FALSE, message = FALSE, warning = FALSE, fig.dim = c(12,10)}  
# Now custom plot to look for errors in catch weight
if (dat.source !="csv") {
options(scipen=10000)
gg <- ggplot(tr_l, aes(LENGTH))+
  geom_histogram(binwidth = 10)+
  theme_bw()+
  ggtitle("Plot of Length Frequency by Species")

ggplotly(gg)
} else {
    print("The data source is csv, so the TR_L table does not exist!")
}
```

# <u>Table of TR_FISH Variables</u>
```{r echo=FALSE, message = FALSE, warning = FALSE}
data("tr_fish_names")
tr_fish <- tr_fish %>% clean_names(case = "all_caps")
obs_names <- names(tr_fish)
#check for required fields that can't be NULL. Excepted for csv inputs
#is 'OP_ID'

if (dat.source =="csv") 
{
required <- c('YEAR', 'VESSEL', 'SERIAL', 
      "SPECIES", "LENGTH")
} else {
  required <- c("TR_FISH_ID", "OP_ID",  "SPECIES", "LENGTH")
}
missed <- setdiff(required, obs_names)
if (length(missed) > 0)
    stop(print(paste0("The tr_fish data are missing the required field(s) ", missed)))

df <- as.data.frame(CruiseCheckeR::dfSmry(tr_fish))
knitr::kable(df)
```

# <u>Plots of TR_FISH Variables</u>
```{r echo=FALSE, message = FALSE, warning = FALSE, fig.dim = c(12,10)}
library(plotly)
# These are canned plots for all variables.
dfPlot(tr_fish, mcex=1.1, cex = 0.8, mfrow=c(2,2))
```

```{r echo=FALSE, message = FALSE, warning = FALSE, fig.dim = c(12,10)}
# Now custom plot to look for errors in catch weight
options(scipen=10000)
gg <- ggplot(tr_fish, aes(LENGTH))+
  geom_histogram(binwidth = 10)+
  facet_wrap(~SPECIES, scales = "free")+
  theme_bw()+
  ggtitle("Plot of Length Distribution by Species")
ggplotly(gg)
#now do weight
gg <- ggplot(tr_fish, aes(WEIGHT))+
  geom_histogram(binwidth = 10)+
  facet_wrap(~SPECIES, scales = "free")+
  theme_bw()+
  ggtitle("Plot of Weight Distribution by Species")
ggplotly(gg)

gg <- ggplot()+
  geom_point(data = tr_fish, aes(x = LENGTH, y = WEIGHT))+
  facet_wrap(~SPECIES, scales = "free")+
  theme_bw()+
  ggtitle("Plot of Length Versus Weight by Species")
ggplotly(gg)
```
